Modal Logics in the Coq Proof Assistant

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
some small modifications wrt abstract and intro

\begin{abstract} 
This paper describes an embedding of higher-order modal
logics in the \Coq proof assistant. {\Coq}'s capabilities are thus extended in a
minimalistic manner, which is nevertheless sufficient for the formalization
of significant modal proofs. The elegance and convenience of this approach, 
from a user perspective, are illustrated here with the successful formalization of Gödel's ontological argument. 
\end{abstract}

@@

Another recently explored possibility \cite{J23,B9,J22} is the embedding of first-order and higher-order modal logics into classical higher-order logics, for which existing higher-order automated theorem provers \cite{ToDo:Leo,Satallax} exist.
Embedding is flexible, because various modal logics can be easily supported by adding their characteristic axioms. Furthermore, it is relatively simple to implement, because it does not require any modification in the source code of the higher-order prover. The prover can be used as is, and only the input files provided to the prover must be especially encoded. Furthermore, the efficacy and efficiency of the embedding approach has been confirmed in philosophically interesting benchmarks \cite{FormalTheologyRepository}. These qualities make embedding a convenient approach for \emph{fully automated} reasoning.

However, one may wonder whether the embedding approach is adequate for \emph{interactive} reasoning, when the user proves theorems by interacting with a proof assistant. The main goal of this paper is to answer this question positively. The \Coq proof assistant was chosen because of the authors' greater familiarity with the tactic language of this system.

@@

In order to illustrate the use of the implemented embedding, we show here the formalization of Scott's version \cite{Scott} of Gödel's ontological argument for God's existence (in Section \ref{sec:proof}). This proof was chosen mainly for two reasons. Firstly, it requires not only modal operators, but also higher-order quantification. Therefore, it is beyond reach for specialized propositional and first-order theorem provers. Secondly, this argument addresses an ancient problem in Philosophy and Metaphysics, which has nevertheless received a lot of attention in the last 15 years, because of the discovery of the modal collapse \cite{ToDo}. This proof lies in the center of a vast and largely unexplored application domain for automated and interactive theorem provers.

>>>>

\begin{abstract} 
This paper describes an embedding of higher-order modal
logics in the \Coq proof assistant. {\Coq}'s capabilities are thus extended in a
minimalistic manner, which is nevertheless sufficient for the formalization
of significant modal proofs. The elegance, flexibility and convenience of this approach, 
from a user perspective, are illustrated here with the successful formalization of Gödel's ontological argument. 
\end{abstract}

@@

Another recently explored possibility \cite{J23,B9,J22} is the
embedding of first-order and higher-order modal logics into classical
higher-order logics, for which existing higher-order automated theorem
provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
because various modal logics can be easily supported by adding their
characteristic axioms.  The embedding approach can also be easily
adapted to support multiple modalities, and it is possible to replace
constant domain quantifiers by varying or cumulative domain
quantifiers.  Moreover, the approach is relatively simple to
implement, because it does not require any modification in the source
code of the higher-order prover. The prover can be used as is, and
only the input files provided to the prover must be especially
encoded. Furthermore, the efficacy and efficiency of the embedding
approach has been confirmed in philosophically interesting and
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant. The main goal of this paper is
to strudy this question. Our answer is positive. The \Coq proof
assistant was chosen because of the authors' greater familiarity with
the tactic language of this system.

@@

In order to illustrate the use of the implemented embedding, we show
here the formalization of Scott's version \cite{Scott} of Gödel's
ontological argument for God's existence (in Section
\ref{sec:proof}). This proof was chosen mainly for two
reasons. Firstly, it requires not only modal operators, but also
higher-order quantification. Therefore, it is beyond reach of
specialized propositional and first-order (modal) theorem provers. Secondly, this argument addresses an ancient problem in Philosophy and Metaphysics, which has nevertheless received a lot of attention in the last 15 years, because of the discovery of the modal collapse \cite{ToDo}. This proof lies in the center of a vast and largely unexplored application domain for automated and interactive theorem provers.
